---
title: "Coronavirus Analysis"
runtime: shiny
output: 
  flexdashboard::flex_dashboard:
    storyboard: true
    orientation: rows
    # social: ["facebook", "twitter", "linkedin"]
    # source_code: embed
    vertical_layout: fill
---

```{r global, include=FALSE}
# TODO: revise the list of packages
# TODO: consider switching to rBokeh
# TODO: modelling and/or simulation
# TODO: display most affected countries
library(shiny)
# library(rvest)
library(flexdashboard)
library(glue)
library(tidyverse)
library(magrittr)
library(lubridate)
library(rmarkdown)
library(patchwork)
library(ggrepel)
library(nCov2019)

# Set colors
confirmed_color <- "purple"
active_color <- "#1f77b4"
recovered_color <- "forestgreen"
death_color <- "firebrick"

# Theme for plots

theme_set(theme_minimal() +
  theme(
    legend.position = "",
    plot.title = element_text(
      size = 15,
      face = "bold",
      colour = "grey40"
    ),
    plot.subtitle = element_text(
      size = 13,
      face = "bold.italic",
      colour = "grey40"
    ),
    plot.caption = element_text(
      size = 12,
      face = "italic",
      colour = "grey40"
    ),
    strip.text = element_text(
      face = "bold",
      size = 15
    ),
    axis.title = element_text(
      size = 15,
      face = "bold"
    ),
    axis.text = element_text(
      size = 13
    )
  ))

all_ncov <- load_nCov2019(lang = "en") %>%
  .["global"] %>%
  tbl_df()

first_case_date <- all_ncov %>%
  group_by(country) %>%
  summarise(first_conf = min(time))

all_ncov %<>%
  left_join(first_case_date) %>%
  mutate(first_conf = as.numeric(time - first_conf)) %>%
  select(time, first_conf, everything())
```



Sidebar {.sidebar}
=======================================================================

This dashboard shows causualty information about the COVID-19 epidemic. More information on the **About** page. 

Select one or more countries to run the analysis. 

```{r}
selectizeInput("country",
  options = list(maxItems = 6),
  label = strong("Select up to six countries"),
  choices = sort(unique(all_ncov$country)),
  selected = c(
    "Italy", "Spain",
    "United States", "United Kingdom",
    "China", "South Korea"
  ),
  multiple = TRUE
)
```

Select also a range of dates to filter the data. By default, the data range is set from February 01 2020 to yesterday,  `r format(Sys.Date()-1, "%B %d %Y")`. 

```{r}
# TODO: last update above reflects time of server, not of publication.
dateRangeInput("daterange",
  label = strong("Date Range"),
  start = min(all_ncov$time),
  end = max(all_ncov$time)
)


checkboxInput("logaxis",
  label = strong("Use log axis?"),
  value = FALSE
)

checkboxInput("startconf",
  label = strong("Start analysis from first case confirmed?"),
  value = FALSE
)

df <- reactive({
  all_ncov %>%
    filter(
      country %in% input$country,
      time >= input$daterange[1],
      time <= input$daterange[2]
    )
})
```


_Last update_:   
**`r format(Sys.time(), "%B %d %Y at %H:%M GMT")`**. 

Summary
=======================================================================

<p style="color:red">**WORK IN PROGRESS**</p>

Row {data-height=900}
-----------------------------------------------------------------------

### **Countries by number of cases**


```{r, fig.width=16}
t1 <- all_ncov %>%
  filter(time == max(time)) %>%
  top_n(10, wt = cum_confirm) %>%
  ggplot(aes(reorder(country, cum_confirm),
    cum_confirm,
    fill = cum_confirm
  )) +
  geom_bar(stat = "identity", color = "grey60") +
  geom_label(
    aes(
      y = max(cum_confirm),
      label = cum_confirm,
      fontface = "bold"
    ),
    hjust = 0, nudge_y = 5000, fill = "white", color = "grey30"
  ) +
  scale_y_continuous(label = scales::comma) +
  scale_fill_viridis_c(direction = -1) +
  theme(
    axis.title.y = element_blank(),
    axis.text.y = element_text(face = "bold")
  ) +
  coord_flip(ylim = c(NA, max(all_ncov$cum_confirm) * 1.2), expand = TRUE) +
  labs(
    title = "Top 10 countries by number of cases",
    subtitle = glue("Latest update on {Sys.Date()-1}"),
    y = "Cases"
  )

t2 <- all_ncov %>%
  filter(time == max(time)) %>%
  top_n(10, wt = cum_dead) %>%
  ggplot(aes(reorder(country, cum_dead),
    cum_dead,
    fill = cum_dead
  )) +
  geom_bar(stat = "identity", color = "grey60") +
  geom_label(
    aes(
      y = max(cum_dead),
      label = cum_dead,
      fontface = "bold",
    ),
    hjust = 0, nudge_y = 200, fill = "white", color = "grey30"
  ) +
  scale_y_continuous(label = scales::comma) +
  scale_fill_viridis_c(direction = -1) +
  theme(
    axis.title.y = element_blank(),
    axis.text.y = element_text(face = "bold")
  ) +
  coord_flip(ylim = c(NA, max(all_ncov$cum_dead) * 1.2), expand = TRUE) +
  labs(
    title = "Top 10 countries by number of fatalities",
    subtitle = glue("Latest update on {Sys.Date()-1}"),
    y = "Cases"
  )

t1 + t2
```


Row
------------------------------------

### 

```{r}
# all_ncov %>% 
#   filter(time == max(time)) %>% 
#   select(country, cum_confirm, cum_dead) %>% 
#   top_n(10, cum_confirm) %>% 
#   pivot_longer(cum_confirm:cum_dead) %>% 
#   ggplot(aes(name, value, group = country)) + 
#   geom_point(aes(color = country))+
#   geom_line(aes(color = country))+
#   scale_y_log10()+
#   geom_label_repel(aes(label = country, fill = country))+
#   theme(axis.title = element_blank(), 
#         axis.text.y = element_blank(), 
#         panel.grid = element_blank())
```

###

Moar yada 

Accumulated cases
=======================================================================

Row 
-----------------------------------------------------------------------

### confirmed {.value-box}

```{r}
renderValueBox({
  max_conf <- df() %>%
    filter(time == min(input$daterange[2], max(time))) %>%
    summarise(cum_confirm = sum(cum_confirm)) %>%
    pluck("cum_confirm")
  valueBox(
    value = paste(format(max_conf, big.mark = ","), "", sep = " "),
    caption = "Total confirmed cases in the countries selected to date",
    icon = "fas fa-user-md",
    color = confirmed_color
  )
})
```

### all confirmed {.value-box}

```{r}
renderValueBox({
  all_conf <- all_ncov %>%
    filter(time == min(input$daterange[2], max(time))) %>%
    summarise(cum_confirm = sum(cum_confirm)) %>%
    pluck("cum_confirm")
  valueBox(
    value = paste(format(all_conf, big.mark = ","), "", sep = " "),
    caption = "Worldwide confirmed cases to date",
    icon = "fas fa-globe",
    color = "slateblue"
  )
})
```

### death {.value-box}

```{r}
renderValueBox({
  max_dead <- df() %>%
    filter(time == min(input$daterange[2], max(time))) %>%
    summarise(cum_dead = sum(cum_dead)) %>%
    pluck("cum_dead")
  valueBox(
    value = paste(format(max_dead, big.mark = ","), "", sep = " "),
    caption = "Total confirmed deaths in the countries selected to date",
    icon = "fas fa-heart",
    color = death_color
  )
})
```

### all death {.value-box}

```{r}
renderValueBox({
  all_dead <- all_ncov %>%
    filter(time == min(input$daterange[2], max(time))) %>%
    summarise(cum_dead = sum(cum_dead)) %>%
    pluck("cum_dead")
  valueBox(
    value = paste(format(all_dead, big.mark = ","), "", sep = " "),
    caption = "Worldwide confirmed deaths to date",
    icon = "fas fa-skull",
    color = "darkred"
  )
})
```

### recovered {.value-box}

```{r}
renderValueBox({
  max_heal <- df() %>%
    filter(time == min(input$daterange[2], max(time))) %>%
    summarise(cum_heal = sum(cum_heal)) %>%
    pluck("cum_heal")
  valueBox(
    value = paste(format(max_heal, big.mark = ","), "", sep = " "),
    caption = "Total recovered cases in the countries selected to date",
    icon = icon("user-shield", lib = "font-awesome"),
    color = recovered_color
  )
})
```

### all confirmed {.value-box}

```{r}
renderValueBox({
  all_heal <- all_ncov %>%
    filter(time == min(input$daterange[2], max(time))) %>%
    summarise(cum_heal = sum(cum_heal)) %>%
    pluck("cum_heal")
  valueBox(
    value = paste(format(all_heal, big.mark = ","), "", sep = " "),
    caption = "Worldwide recovered cases to date",
    icon = "fas procedures",
    color = "darkgreen"
  )
})
```


Row {.tabset}
-----------------------------------------------------------------------

### **Cumulative cases per country**

```{r}
renderPlot({
  cumplot <- df() %>%
    mutate(end_label = ifelse(time == max(time), glue("{country}: {scales::comma(cum_confirm)}"), NA))

  if (input$startconf) {
    cumplot %<>% ggplot(aes(first_conf, cum_confirm, color = country))
  } else {
    cumplot %<>% ggplot(aes(time, cum_confirm, color = country))
  }

  cumplot <- cumplot +
    geom_line() +
    geom_label_repel(aes(label = end_label, fill = country),
      nudge_x = .1,
      nudge_y = .5,
      size = 5,
      color = "white",
      show.legend = FALSE
    ) +
    labs(caption = glue("Data last updated on {max(df()$time)}"))

  if (input$startconf) {
    cumplot <- cumplot + labs(x = "Days elapsed since first confirmed case")
  } else {
    cumplot <- cumplot + labs(x = "Date")
  }


  if (input$logaxis) {
    cumplot + scale_y_log10("Log number of cases", labels = scales::comma)
  } else {
    cumplot + scale_y_continuous("Number of cases", labels = scales::comma)
  }
})
```

### **Cumulative deaths per country**

```{r}
renderPlot({
  deathplot <- df() %>%
    mutate(end_label = ifelse(time == max(time), glue("{country}: {scales::comma(cum_dead)}"), NA))

  if (input$startconf) {
    deathplot %<>% ggplot(aes(first_conf, cum_dead, color = country))
  } else {
    deathplot %<>% ggplot(aes(time, cum_dead, color = country))
  }


  deathplot <- deathplot +
    geom_line() +
    geom_label_repel(aes(label = end_label, fill = country),
      nudge_x = .1,
      nudge_y = .5,
      size = 5,
      color = "white",
      show.legend = FALSE
    ) +
    labs(caption = glue("Data last updated on {max(df()$time)}"))

  if (input$startconf) {
    deathplot <- deathplot + labs(x = "Days elapsed since first confirmed case")
  } else {
    deathplot <- deathplot + labs(x = "Date")
  }

  if (input$logaxis) {
    deathplot + scale_y_log10("Log number of deaths", labels = scales::comma)
  } else {
    deathplot + scale_y_continuous("Number of deaths", labels = scales::comma)
  }
})
```

### **Cumulative recoveries per country**

```{r}
renderPlot({
  recplot <- df() %>%
    mutate(end_label = ifelse(time == max(time), glue("{country}: {scales::comma(cum_heal)}"), NA))

  if (input$startconf) {
    recplot %<>% ggplot(aes(first_conf, cum_heal, color = country))
  } else {
    recplot %<>% ggplot(aes(time, cum_heal, color = country))
  }

  recplot <- recplot +
    geom_line() +
    geom_label_repel(aes(label = end_label, fill = country),
      nudge_x = .1,
      nudge_y = .5,
      size = 5,
      color = "white",
      show.legend = FALSE
    ) +
    labs(caption = glue("Data last updated on {max(df()$time)}"))

  if (input$startconf) {
    recplot <- recplot + labs(x = "Days elapsed since first confirmed case")
  } else {
    recplot <- recplot + labs(x = "Date")
  }

  if (input$logaxis) {
    recplot + scale_y_log10("Log number of recovered", labels = scales::comma)
  } else {
    recplot + scale_y_continuous("Number of recovered", labels = scales::comma)
  }
})
```

Progression
===========================================================================

Row {data-height=750}
---------------------------------------------------------------------------

### **Acceleration of new cases**

```{r}
# TODO: Add valueboxes on top with number of daily cases
# TODO: Refactor as a single function

df_speed <- reactive({
  df() %>%
    arrange(country, time) %>%
    group_by(country) %>%
    mutate(
      cases_confirm = cum_confirm - lag(cum_confirm),
      cases_recover = cum_heal - lag(cum_heal),
      cases_dead = cum_dead - lag(cum_dead)
    ) %>%
    filter(!is.na(cases_confirm))
})

renderPlot({
  d <- df_speed() %>%
    mutate(seconddif = cases_confirm - lag(cases_confirm))
  dmin <- d %>%
    filter(time == min(time))

  ggplot(d) +
    aes(time, seconddif) +
    geom_rect(
      data = dmin,
      aes(
        xmin = time,
        xmax = input$daterange[2],
        ymin = -Inf, ymax = 0
      ),
      fill = "forestgreen",
      alpha = .3
    ) +
    geom_rect(
      data = dmin,
      aes(
        xmin = time,
        xmax = input$daterange[2],
        ymin = 0, ymax = Inf
      ),
      alpha = .3,
      fill = "firebrick"
    ) +
    scale_x_date(date_breaks = "1 month", date_labels = "%b %d") +
    scale_y_continuous(labels = scales::comma) +
    facet_wrap(~country, scales = "free") +
    geom_hline(yintercept = 0, color = "grey50", linetype = 2) +
    geom_line(aes(color = country), alpha = .8) +
    geom_smooth(aes(color = country),
      method = "loess", span = 1 / 3,
      se = FALSE, size = 1
    ) +
    labs(
      x = "",
      y = "Diff cases",
      title = "New cases acceleration rate",
      caption = "Y axes not to scale"
    ) +
    theme(panel.grid = element_blank())
})
```

### **Acceleration of new fatalities** 

```{r}
renderPlot({
  d <- df_speed() %>%
    mutate(seconddif = cases_dead - lag(cases_dead))
  dmin <- d %>%
    filter(time == min(time))

  ggplot(d) +
    aes(time, seconddif) +
    scale_y_continuous(labels = scales::comma) +
    scale_x_date(date_breaks = "1 month", date_labels = "%b %d") +
    geom_rect(
      data = dmin,
      aes(
        xmin = time,
        xmax = input$daterange[2],
        ymin = -Inf, ymax = 0
      ),
      fill = "forestgreen",
      alpha = .3
    ) +
    geom_rect(
      data = dmin,
      aes(
        xmin = time,
        xmax = input$daterange[2],
        ymin = 0, ymax = Inf
      ),
      alpha = .3,
      fill = "firebrick"
    ) +
    facet_wrap(~country, scales = "free") +
    geom_hline(yintercept = 0, color = "grey50", linetype = 2) +
    geom_line(aes(color = country), alpha = .8) +
    geom_smooth(aes(color = country),
      method = "loess", span = 1 / 3,
      se = FALSE, size = 1
    ) +
    labs(
      x = "",
      y = "Diff cases",
      title = "New fatalities acceleration rate",
      caption = "Y axes not to scale"
    ) +
    theme(panel.grid = element_blank())
})
```


Row 
---------------------------------------------------------------------------

### **Progression of COVID-19**

Speed of the disease can be defined as the rate in which new cases or new fatalities appear. In this sense, **speed** is just the number of new cases reported per day. A speed of zero means no new cases ocurred. In this context, speed cannot be below zero, as it would be meaningless (although sometimes a negative number is reported to correct previously inflated cases, a general bad practice). 

**Acceleration** is therefore the _second order difference_, that is, the difference between today's speed and yesterday's. An acceleration of _zero_ means that today's increase is the same as yesterday's, and an _negative_ acceleration means that the rate at which new cases happen is reducing. A _positive_ acceleration, on the other hand, means that the rate of new cases is increasing - today's new cases is larger than yesterday's. 

To sum up, we can envision four scenarios (from better to worse): 

1. __Speed = 0__ and __Acceleration = 0__:  best case scenario, no new cases today same as yesterday. 
1. __Speed > 0__ and __Acceleration < 0__:  New cases appeared, but less than yesterday. 
1. __Speed > 0__ and __Acceleration = 0__: New cases appeared, at the same rate as yesterday. 
1. __Speed > 0__ and __Acceleration > 0__: New cases appeared, and at a higher rate than yesterday. 


To smooth over the daily noise in reporting these numbers, we can take a moving average (loess). 


About
=======================================================================

**The Coronavirus Dashboard**

This dashboard summarises the status of the 2019-2020 coronavirus epidemic COVID-19 in the world. The dashboard is build using the  [flexdashboards](https://rmarkdown.rstudio.com/flexdashboard/index.html){target="_blank"} by RStudio and published to [shinyapps.io](https://www.shinyapps.io/){target="_blank"}. 


**Code**

The code behind this dashboard is available on [GitHub](https://github.com/Lornebradia/dashboards/blob/master/dashboards/coronavirus.Rmd){target="_blank"}.

**Data**

The data used for this report comes from a variety of sources, gathered in the [`nCov2019` R package](https://github.com/GuangchuangYu/nCov2019){target="_blank"}. This package provides up-to date information about the COVID-19 epidemic. The dataset focuses on China, but it also contains functions to access data from the rest of the world. The package [vignette](https://guangchuangyu.github.io/nCov2019/){target="_blank"} explains the general functionality of the package.

You can install this package by running: 

```
remotes::install_github("GuangchuangYu/nCov2019")
```

The data and dashboard are refreshed on a daily basis.

**Contact**

Comments & suggestions welcome! You can [open an issue](https://github.com/Lornebradia/dashboards/issues/new){target="_blank"} in the repository for the dashboard. 
